---
title: "PPOL 628 Final Project"
author: "Dain Yoo"
date: "12/14/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse) 
library(tidymodels)
library(ggplot2)
library(gridExtra)
library(grid)
library(tidyr)
library(GGally)
library(ggpubr)
library(factoextra)
library(viridis)
```

### Research puzzle

I am interested in how Egyptian public opinion shifts after the adoption of anti-terror law in Egypt. Specifically, I am interested in exploring polarization between secularists and Islamists as anti-terror law, by nature, targets Islamist extremists and the secular regime would target the Islamists as potential threat to national security. 


### Data and Methods

I use Arab Barometer survey data (2006-2019) to observe latent clusters within public opinion. Because Egypt adopted anti-terror in 2015, I use Wave 3 (2012-2014) and Wave 4 (2016-2017) to see how polarization exacerbated in Egypt after the enactment. 

Among various questionnaires, I includeㅇ 20 of them for more variation across clusters. They include binary variables (e.g. urban/rural, male/female, voted/did not vote), categorical ordinal variable (e.g. not at all/once a month/once a week/daily, level of education) and continuous variable (e.g. age). Topics of the selected questionnaires include political participation, religious practices, media exposure and stances on economic and diplomatic issues.


```{r, warning=FALSE}
# Read survey dataset
abv3 <- read_csv("abv3.csv")
abv4 <- read_csv("abv4.csv")

# Subset for Egypt and 20 survey questions
egypt1 <- abv3 %>%
  filter(country == "Egypt") %>%
  select(country, wt, q13, q1002, q1012, q1004, q1001, q1003,
         q102, q105, q106, q20112, q210, q301, q404, q409, q513,
         q523, q605a, q609, q6105, q6101)

egypt2 <- abv4 %>%
  filter(country == "Egypt") %>%
  select(country, wt, q13, q1002, q1012, q1004, q1001, q1003,
         q102, q105, q106, q1072, q210, q301, q404, q409, q513,
         q523, q605a, q609, q6106, q6101)
```


### Preprocessing - Survey weight

To preprocess the data before analyzing, I account for the weight of the survey to get an ideal representation of the population. As for the first dataset (2012-2014), the weights range from 0.047 to 6.07 while that of the second dataset (2016-2017) range from 0.67 to 1.90. To keep the number of observations similar, I multiply the weights by 10 and round to a whole number. Then I duplicate the each observation by the rounded weight. 

```{r}
# Account for weights
summary(egypt1$wt)
summary(egypt2$wt)


# Duplicated observations by weight*10
# and rounded to a whole number
wt1 <- egypt1 %>%
  mutate(wt10 = wt*10, wt_round = round(wt10, 0))
wt2 <- egypt2 %>%
  mutate(wt10 = wt*10, wt_round = round(wt10, 0))

egypt_wt1 <- wt1 %>%
  uncount(wt_round)
egypt_wt2 <- wt2 %>%
  uncount(wt_round)
```


```{r}
## Descriptive Plot
# Egypt
ggplot(data = tun, aes(x = "", y = n, fill = q20112)) + 
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  coord_polar("y", start = 0) +
  scale_fill_brewer(palette="Blues") +
  facet_grid(facets=. ~ year) + theme_bw() + 
  labs(title = "Attitude towards Islamist Parties in Tunisia",
       subtitle = "Before and After Anti-Terror Law Adoption in 2015",
       caption = "Source: Arab Barometer Wave III, IV") + 
  theme(legend.title = element_blank(),
        panel.grid  = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())


```


## Normalize responses



```{r}
# Recode responses (wave 3)
egypt_1 <- egypt_wt1 %>%
  mutate(q1001 = as.numeric(q1001),
         q1001 = q1001/50) %>%
  mutate(q13 = recode(q13,
                        "Rural" = 0,
                        "Urban" = 1)) %>%
  mutate(q1002 = recode(q1002,
                        "Female" = 0,
                        "Male" = 1)) %>%  
  mutate(q1012 = recode(q1012,
                        "Christian" = 0,
                        "Muslim" = 1)) %>%
  mutate(q1004 = recode(q1004,
                        "No" = 0,
                        "Yes" = 1)) %>%
    filter(!q301 %in% c("Refuse", 
                        "Don't know")) %>%
    mutate(q301 = recode(q301,
                        "No" = 0,
                        "Yes" = 1)) %>%
    filter(!q210 %in% c("Don't know")) %>%
    mutate(q210 = recode(q210,
                        "No" = 0,
                        "Yes" = 1)) %>%
    mutate(q1003 = recode(q1003,
                          "Illiterate/No formal education" = 0.07,
                          "Elementary" = 0.21,
                          "Prepartory/Basic" = 0.36,
                          "Secondary" = 0.5,
                          "Mid-level diploma (professional or technical" = 0.64,
                          "ba" = 0.79,
                          "MA and above" = 0.96)) %>%
  filter(!q6101 %in% c("Refuse", 
                        "Don't know")) %>%
  mutate(q6101 = recode(q6101,
                        "Never" = 0.1,
                        "Rarely" = 0.3,
                        "Sometimes" = 0.5,
                        "Most of the time" = 0.7,
                        "Always"= 0.9)) %>%
  filter(!q6105 %in% c("Refuse", 
                        "Don't know")) %>%
  mutate(q6105 = recode(q6105,
                        "Never" = 0.1,
                        "Rarely" = 0.3,
                        "Sometimes" = 0.5,
                        "Most of the time" = 0.7,
                        "Always"= 0.9)) %>%
  filter(!q609 %in% c("Refuse")) %>%
  mutate(q609 = recode(q609,
                        "Not religious" = 0.17,
                        "Somewhat religious" = 0.5,
                        "Religious" = 0.83)) %>%
  filter(!q605a %in% c("Don't know")) %>%
  mutate(q605a = recode(q605a,
                        "I Strongly agree with the 1st sentence" = 0.1,
                        "I agree with the 1st sentence" = 0.3,
                        "I do not agree with either sentence (Do not read)" = 0.5,
                        "I agree with the 2nd sentence" = 0.7,
                        "I strongly agree with the 2nd sentence"= 0.9)) %>%
  filter(!q523 %in% c("Don't know")) %>%
  mutate(q523 = recode(q523,
                        "Not justified at all" = 0.125,
                        "Justified to a limited extent" = 0.375,
                        "Justified to a medium extent" = 0.625,
                        "Justified to a great extent" = 0.875)) %>%
  filter(!q513 %in% c("Don't know",
                      "Refuse",
                      "Not concerned / Not interested")) %>%
  mutate(q513 = recode(q513,
                       "1" = 1,
                       "2" = 2,
                       "3" = 3,
                       "4" = 4,
                       "5" = 5, 
                       "6" = 6,
                       "7" = 7,
                       "8" = 8,
                       "9" = 9,
                        "Absolutely unsatisfied" = 0.1,
                        "Completely satisfied" = 10),
         q513 = q513/10) %>%
  filter(!q409 %in% c("Don't know")) %>%
  mutate(q409 = recode(q409,
                         "Daily or almost daily" = 0.1 ,
                         "At least once a week" = 0.3,
                         "At least once a month" = 0.5,
                         "A few times a year" = 0.7,
                         "I do not use the internet" = 0.9)) %>%
  filter(!q404 %in% c("Don't know")) %>%
  mutate(q404 = recode(q404,
                         "Very interested" = 0.125,
                         "Interested" = 0.375,
                         "Slightly interested" = 0.625,
                         "Not interested" = 0.875)) %>%
  filter(!q301 %in% c("Refuse", 
                      "Don't know")) %>%  
  filter(!q210 %in% c("Don't know")) %>%
  filter(!q20112 %in% c("Refuse", 
                        "Don't know")) %>% 
  mutate(q20112 = recode(q20112,
                         "I absolutely do not trust it" = 0.125,
                         "I trust it to a limited extent" = 0.375,
                         "I trust it to a medium extent" = 0.625,
                         "I trust it to a great extent" = 0.875)) %>%
  filter(!q106 %in% c("Refuse", 
                      "Don't know")) %>% 
  mutate(q106 = recode(q106,
                         "Not at all" = 0.125,
                         "To a limited extent" = 0.375,
                         "To a medium extent" = 0.625,
                         "To a great extent" = 0.875)) %>%
  filter(!q105 %in% c("Refuse", 
                      "Don't know")) %>% 
  mutate(q105 = recode(q105,
                         "Absolutely not ensured" = 0.125,
                         "Not ensured" = 0.375,
                         "Ensured" = 0.625,
                         "Fully ensured" = 0.875)) %>%  
  filter(!q102 %in% c("Refuse", 
                      "Don't know")) %>% 
  mutate(q102 = recode(q102,
                         "Much worse" = 0.1,
                         "Somewhat worse" = 0.3,
                         "Almost the same as the current situation" = 0.5,
                         "Somewhat better" = 0.7,
                         "Much better" = 0.9))
```


```{r}
# Recode responses (wave 4)
egypt_2 <- egypt_wt2 %>%
  mutate(q1001 = as.numeric(q1001),
         q1001 = q1001/50) %>%
  mutate(q13 = recode(q13,
                        "Rural" = 0,
                        "Urban" = 1)) %>%
  mutate(q1002 = recode(q1002,
                        "Female" = 0,
                        "Male" = 1)) %>%  
  mutate(q1012 = recode(q1012,
                        "Christian" = 0,
                        "Muslim" = 1)) %>%
  mutate(q1004 = recode(q1004,
                        "No" = 0,
                        "Yes" = 1)) %>%
    filter(!q301 %in% c("Decline to answer (Do not read)")) %>%
    mutate(q301 = recode(q301,
                        "No" = 0,
                        "Yes" = 1)) %>%
    mutate(q1003 = recode(q1003,
                          "Illiterate/No formal education" = 0.07,
                          "Elementary" = 0.21,
                          "Preparatory/Basic" = 0.36,
                          "Secondary" = 0.5,
                          "Mid-level diploma/professional or technical" = 0.64,
                          "BA" = 0.79,
                          "MA and above" = 0.96)) %>%
  filter(!q6101 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>%
  mutate(q6101 = recode(q6101,
                        "Never" = 0.1,
                        "Rarely" = 0.3,
                        "Sometimes" = 0.5,
                        "Most of the time" = 0.7,
                        "Always"= 0.9)) %>%
  filter(!q6106 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>%
  mutate(q6106 = recode(q6106,
                        "Never" = 0.1,
                        "Rarely" = 0.3,
                        "Sometimes" = 0.5,
                        "Most of the time" = 0.7,
                        "Always"= 0.9)) %>%
  filter(!q609 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>%
  mutate(q609 = recode(q609,
                        "Not religious" = 0.17,
                        "Somewhat religious" = 0.5,
                        "Religious" = 0.83)) %>%
  filter(!q605a %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>%
  mutate(q605a = recode(q605a,
                        "I strongly agree with the 1st sentence" = 0.1,
                        "I agree with the 1st sentence" = 0.3,
                        "Do not agree with either (Do not read)" = 0.5,
                        "I agree with the 2nd sentence" = 0.7,
                        "I strongly agree with the 2nd sentence"= 0.9)) %>%
  filter(!q523 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>%
  mutate(q523 = recode(q523,
                        "Not justifiable at all" = 0.125,
                        "Not very justified" = 0.375,
                        "Somewhat justified" = 0.625,
                        "Completely justified" = 0.875)) %>%
  filter(!q513 %in% c("Decline to answer (Do not read)",
                      "Don't know (Do not read)",
                      "Not concerned/not interested")) %>%
  mutate(q513 = recode(q513,
                       "1" = 1,
                       "2" = 2,
                       "3" = 3,
                       "4" = 4,
                       "5" = 5, 
                       "6" = 6,
                       "7" = 7,
                       "8" = 8,
                       "9" = 9,
                       "Completely unsatisfied" = 0.1,
                      "Completely satisfied" = 10),
         q513 = q513/10) %>%
  filter(!q409 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>%
  mutate(q409 = recode(q409,
                         "I do not use the internet " = 0.083 ,
                         "Less than once a week" = 0.25,
                         "Once a week" = 0.417,
                         "Several times a week" = 0.583,
                         "Daily" = 0.75,
                         "I am online almost all day" = 0.91)) %>%
  filter(!q404 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>%
  mutate(q404 = recode(q404,
                         "Very interested" = 0.125,
                         "Interested" = 0.375,
                         "Somewhat interested" = 0.625,
                         "Not interested at all" = 0.875)) %>%
  filter(!q301 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>%  
  filter(!q210 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>%
  mutate(q210 = recode(q210,
                         "Not at all" = 0.125,
                         "To a small extent" = 0.375,
                         "To a medium extent" = 0.625,
                         "To a large extent" = 0.875)) %>%
  filter(!q1072 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>% 
  mutate(q1072 = recode(q1072,
                         "Not at all" = 0.125,
                         "Not much" = 0.375,
                         "Much" = 0.625,
                         "Very much" = 0.875)) %>%
  filter(!q106 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>% 
  mutate(q106 = recode(q106,
                         "Not treated equally at all" = 0.125,
                         "To a limited extent" = 0.375,
                         "To a medium extent" = 0.625,
                         "To a great extent" = 0.875)) %>%
  filter(!q105 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>% 
  mutate(q105 = recode(q105,
                         "Absolutely not ensured" = 0.125,
                         "Not ensured" = 0.375,
                         "Ensured" = 0.625,
                         "Fully ensured" = 0.875)) %>%  
  filter(!q102 %in% c("Decline to answer (Do not read)", 
                        "Don't know (Do not read)")) %>% 
  mutate(q102 = recode(q102,
                         "Much worse" = 0.1,
                         "Somewhat worse" = 0.3,
                         "Almost the same as the current situation" = 0.5,
                         "Somewhat better" = 0.7,
                         "Much better" = 0.9))
```    



```{r}
# Drop weight variables
egypt_1 <- select(egypt_1, -one_of(c("wt", "wt10", "country")))
egypt_2 <- select(egypt_2, -one_of(c("wt", "wt10", "country"))) 


#Optimal number of clusters
fviz_nbclust(egypt_1, FUN = kmeans, method = "wss") #5?
fviz_nbclust(egypt_1, FUN = kmeans, method = "silhouette") #2
fviz_nbclust(egypt_1, FUN = kmeans, method = "gap_stat") #7


fviz_nbclust(egypt_2, FUN = kmeans, method = "wss") #5?
fviz_nbclust(egypt_2, FUN = kmeans, method = "silhouette") #2
fviz_nbclust(egypt_2, FUN = kmeans, method = "gap_stat") #9

#Clustering Analysis
set.seed(2020)
egypt1_kmeans <- kmeans(egypt_1,
                       centers = 6, # number of clusters
                       nstart = 100)
egypt2_kmeans <- kmeans(egypt_2,
                       centers = 6, # number of clusters
                       nstart = 100)

tidy(egypt1_kmeans) %>% knitr::kable(digits = 2, "simple")
tidy(egypt2_kmeans) %>% knitr::kable(digits = 2, "simple")

#Expert check
mean1 <- bind_cols(
  egypt_1,
cluster = egypt1_kmeans$cluster) %>%
  group_by(cluster) %>% 
  summarize(q13 = mean(q13),
            q1002 = mean(q1002),
            q1012 = mean(q1012),
            q1004 = mean(q1004),
            q1001 = mean(q1001),
            q1003 = mean(q1003),
            q102 = mean(q102),
            q105 = mean(q105),
            q106 = mean(q106),
            q20112 = mean(q20112),
            q210 = mean(q210),
            q301 = mean(q301),
            q404 = mean(q404),
            q409 = mean(q409),
            q513 = mean(q513),
            q523 = mean(q523),
            q605a = mean(q605a),
            q609 = mean(q609),
            q6105 = mean(q6105),
            q6101 = mean(q6101)) 

mean1 %>% knitr::kable(digits = 3)

mean2 <- bind_cols(
  egypt_2,
  cluster = egypt2_kmeans$cluster) %>%
  group_by(cluster) %>% 
  summarize(q13 = mean(q13),
            q1002 = mean(q1002),
            q1012 = mean(q1012),
            q1004 = mean(q1004),
            q1001 = mean(q1001),
            q1003 = mean(q1003),
            q102 = mean(q102),
            q105 = mean(q105),
            q106 = mean(q106),
            q1072 = mean(q1072),
            q210 = mean(q210),
            q301 = mean(q301),
            q404 = mean(q404),
            q409 = mean(q409),
            q513 = mean(q513),
            q523 = mean(q523),
            q605a = mean(q605a),
            q609 = mean(q609),
            q6106 = mean(q6106),
            q6101 = mean(q6101)) 

mean2 %>%  knitr::kable(digits = 3)
```



```{r}
#PCA
principle_components1 <- prcomp(egypt_1, scale. = FALSE)
principle_components2 <- prcomp(egypt_2, scale. = FALSE)

# view the amount of variance explained
summary(principle_components1)

# create and visualize the group membership of observations in the components
bind_cols(
  as_tibble(principle_components1$x), 
  cluster = egypt1_kmeans$cluster) %>%
  ggplot(aes(PC1, PC2, color = factor(cluster))) + geom_point() +
  labs(title = "Public Opinion Clusters and PCA") + theme_minimal()

bind_cols(
  as_tibble(principle_components2$x), 
  cluster = egypt2_kmeans$cluster) %>%
  ggplot(aes(PC1, PC2, color = factor(cluster))) + geom_point() +
  labs(title = "Public Opinion Clusters and PCA") + theme_minimal()
```

```{r, include= FALSE}
#change as factor
mean1$cluster <- as.factor(mean1$cluster)
mean2$cluster <- as.factor(mean2$cluster)
```


```{r}
#Parellel coordinate plot
x <- ggparcoord(mean1, columns = c(2,3,5:7,11,14:19), 
                groupColumn = 'cluster', scale = 'globalminmax',
                showPoints = TRUE) + theme_bw() + scale_color_viridis(discrete=TRUE)

y <- ggparcoord(mean2, columns = c(2,3,5:7,11,14:19), 
                groupColumn = 'cluster', scale = 'globalminmax',
                showPoints = TRUE) + theme_bw() + scale_color_viridis(discrete=TRUE)

ggarrange(
  x, y, labels = c("T1", "T2"),
  common.legend = TRUE, legend = "bottom", nrow = 2
  )
```


